\chapter{Binary Encoding Using Dichotomies}

Dichotomies can be used to find a binary encoding for a state machine. A dichotomy is a pair of sets called $L$ and $R$. Both sets can contain an arbitrary number of symbols. In the case of state machines, a symbol represents a state.

The first step necessary for binary encoding using dichotomies is to generate a constraint matrix $A$. It is derived from a minimal cover of the state machine. The constraint matrix contains all combined states that combine multiple states, but not all states of the state machine.

Before explaining the algorithm to find a binary encoding, some terms have to be defined:
\begin{itemize}
\item \textbf{Root dichotomy of a row $a^T$:} A root dichotomy of a row $a^T$ of $A$ is a dichotomy where $L$ contains all symbols that have a 1 in $a^T$ and $R$ contains one symbol that has a 0 in $a^T$.
\item \textbf{Compatibility:} $(L_1,R_1)$ and $(L_2,R_2)$ are compatible if one of the following statements holds true:
\begin{itemize}
\item $L_1 \cap R_2 = \emptyset$ and $R_1 \cap L_2 = \emptyset$
\item $L_1 \cap L_2 = \emptyset$ and $R_1 \cap R_2 = \emptyset$
\end{itemize}
\item \textbf{Coverage:} $(L_1,R_1)$ covers $(L_2,R_2)$ if one of the following statements holds true:
\begin{itemize}
\item $L_1 \supseteq R_2$ and $R_1 \supseteq L_2$
\item $L_1 \supseteq L_2$ and $R_1 \supseteq R_2$
\end{itemize}
\item \textbf{Prime dichotomy:} A prime dichotomy is a dichotomy that can not be covered by a compatible dichotomy.
\end{itemize}

The algorithm to find an exact binary encoding works in the following way:
\begin{enumerate}
\item Compute all prime dichotomies.
\item Compute all root dichotomies for the constraint matrix $A$.
\item Generate a table with one column for every root dichotomy and one row for every prime dichotomy. Every cell of the table is set to 1 if the prime dichotomy covers the root dichotomy. Otherwise, the cell is set to 0.
\item Find a minimal set of prime dichotomies that cover all root dichotomies.
\item Generate the encoding matrix from the $L$ set of the found prime dichotomies.
\end{enumerate}

\section{Task}

The task given was to implement exact binary encoding using dichotomies. A framework was provided that can read \texttt{.kiss} files. Those \texttt{.kiss} files describe a state machine. The framework has to be extended by the algorithms to generate the dichotomies and to find a binary encoding of all states. The results should be written to a \texttt{.blif} file that can then be read by \texttt{abc} to map the design to lookup tables. Finally an evaluation should be done to assess the performance of the program.